---
title: "Analyzing changes in CO2 by Sector"
author: "Uni Lee"
date: "1/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(janitor)
library(data.table)
library(ggthemes)

```

# Time to decarbonize our transports!

For years, I thought renewable energy was the "silver bullet" to solving the climate crisis. I campaigned for renewable energy for years at all kinds of organizations - Greenpeace, KFEM and small social enterprises promoting renewable energy solutions for developing countries. Renewable energy was my thing. 

I knew in the back of my head that there was a lot of work to be done in other sectors, too. But the energy world fascinated me so much that I wasn't really looking into other high-carbon sectors like the transport.

What fascinated me about the energy sector was that there was a very clear target to and call to action: the government must act. There is a big role for public policy to play in the energy sector because change depends on the decision of the government or utility companies (which are heavily regulated). There is a smaller role for individual action, because consumers don't even have any option to choose from. There is literally no special "green plug" to buy when people turn their lights on or plug in their electric devices ([natural monopoly](https://cs.stanford.edu/people/eroberts/cs201/projects/corporate-monopolies/benefits_natural.html#)). In contrast, in other sectors like transportation or agriculture, people have to make conscious choices every day: _should I buy an electric or diesel car? Should I buy organic or conventional?_ There is a big role for individuals to play to create social change. Since I prefer living in a free society myself, I don't feel that I have the right to convince people to change their behavior. Besides, I don't want to judge others' consumption or lifestyle choices because there may be justifiable reasons that I am not aware of - not having enough money, for example. And this thinking led me to turn my interest away from other sectors.  

But. [This blog](https://energyathaas.wordpress.com/2021/01/11/a-tale-of-two-sectors/) from the Energy at Haas convinced me that it is indeed about time to change the transport sector. The author, Catherine Wolfram, argues that the progress on carbon emissions reduction in the transportation sector is much more sluggish than the electricity sector in UK, California and Texas. I was curious to see if this is true for other countries and at the global level. 

## Analysis of historical emissions data from Climate Watch 

I extended her analysis to the global level and other countries using the same dataset that she used: [Climate Watch Historical Country Greenhouse Gas Emissions Data (1990-2017)](https://www.climatewatchdata.org/data-explorer/historical-emissions?historical-emissions-data-sources=cait&historical-emissions-gases=all-ghg&historical-emissions-regions=All%20Selected&historical-emissions-sectors=total-including-lucf&page=1). The dataset contains sector‐level greenhouse gas (GHG) emissions data at the global level and country-level. 


### Data Processing

Sectors in this dataset are categorized into six main sectors (Agriculture, Bunker Fuels, Energy, Industrial Processes, Land-Use Change and Forestry, Waste). The "energy" category is further broken down into six sub-sectors: building, electricity/heat, fugitive emissions, manufacturing/construction, fuel combustion and transportation. So, I had to take out the value for the energy category to prevent double-counting. 

```{r}
# Load data and make sure that the data formats are appropriate. Pivot the dataset longer to create "year" column. 
cait_raw <- read.csv("data/historical_emissions.csv", sep=";", stringsAsFactors=FALSE) %>% clean_names()
cait_raw[] <- lapply(cait_raw, function(x) gsub("N/A", NA, x))
cait_raw[,6:33] <- as.numeric(unlist(cait_raw[,6:33]))
cait_raw <- cait_raw %>% pivot_longer(cols=starts_with("x"), names_to="year", values_to="value", names_prefix="x") %>% mutate(year=as.numeric(year))

# Make a subset for sub-sector level values. Filter out the energy category and total values, since the level of granularity is higher. Mark energy sub-sectors with a prefix "Energy". 
category <- read.csv("data/sector_category.csv", sep=";") 
energy_sub <- category$subsector_1[category$sector=="Energy"]

cait_subsector <- cait_raw %>% mutate(sector = ifelse(sector %in% energy_sub, paste("Energy:", sector, sep=" "), sector)) %>% filter(!str_detect(sector, "^Total")) %>% filter(sector != "Energy") %>% na.omit()

```

For the purpose of this analysis, I will consider the relative change in GHG levels compared to 1990s for the year 2000-2017.

```{r}
# Create a new column for the base level (GHG level for 2000s). # Calculate the relative change in GHG compared to the base value. 
 
base <- cait_subsector %>% filter(year==2000) %>% select(country, sector, base = value) %>% na.omit()
cait_ch <- inner_join(cait_subsector,  base, by=c("country", "sector")) %>% mutate(change = round((value-base)/base*100,2)) %>% filter(year>= 2000)
```

For regional-level analysis, calculated the change in total GHG since 2000s by region and OECD membership. Since the original dataset does not contain country's meta data, I sourced another dataset provided by [World Integrated Trade Solution](https://wits.worldbank.org/CountryProfile/Metadata/en/Country/All).

```{r}
# Add country metadata
meta <- read.csv("data/WITSCountryProfile.csv", sep=",") %>% clean_names() %>% mutate(OECD = ifelse(other_groups=="OECD", T, F)) %>% select(country=country_name, region, OECD) 
cait_meta <- left_join(cait_ch, meta, by="country") %>% filter(region!=is.na(region)) %>% filter(region!="") %>% select(-base, -change)

# Calculate the total GHG and their change since 2000 by region
region <- cait_meta %>% group_by(region, sector, year) %>% summarise(value=sum(value, rm.na=T)) 
base_reg <-  region %>% filter(year==2000) %>% select(region, sector, base = value) 
region_ch <- inner_join(region,  base_reg, by=c("region", "sector")) %>% mutate(change = round((value-base)/base*100,2)) %>% filter(year>= 2000)

# Calculate the total GHG and their change in 2000 by OECD membership
oecd <- cait_meta %>% group_by(OECD, sector, year) %>% summarise(value=sum(value, rm.na=T)) 
base_oecd <- oecd %>% filter(year==2000) %>% select(OECD, sector, base = value) 
oecd_ch <- inner_join(oecd,  base_oecd, by=c("OECD", "sector")) %>% mutate(change = round((value-base)/base*100,2)) %>% filter(year>= 2000) %>% mutate(country=ifelse(OECD==TRUE, "OECD", "non-OECD")) %>% ungroup() %>% select(-OECD)

```

### Top 5 GHG emitters have decreased GHG in the electricity sector more than transportation.

```{r include=FALSE}
# Take the names of top 6 emitting countries
cait_top <- cait_raw %>% filter(country!="World" & year==2017) %>% filter(sector=="Total including LUCF") %>% arrange(desc(value)) 
top5 <- paste(cait_top$country[1:6], sep=",", collapse=", ")

```

Catherine's analysis holds true for top 6 emitters: `r top5`

```{r}
# Change in GHG in top 6 emitting countries, 2000-2017.
entr <- c("Energy: Electricity/Heat", "Energy: Transportation")
top5_entr <- cait_ch %>% filter(country %in% cait_top$country[1:6]) %>% filter(sector %in% entr) %>% mutate(country=factor(country, levels=unique(country))) # This is for creating order in the visualization.

ggplot(top5_entr, aes(x=year, y=change, color=sector)) + geom_line() +
  facet_wrap(~country) + theme_stata() + theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title = "Change in GHG in top 6 emitting countries compared to 2000 level", x="Year", y="%") + scale_color_manual(values = c("#20639B", "#ED553B"))

```

Why? => write about the uptake of renewables in China and India. 

### At the global level, GHG reduction in the electricity is still lagging.  

```{r}
# Change in GHG from electricity/heat vs transport compared to 1990 level (1990-2017) - Global vs. OECD vs. non_OECD

world <- cait_ch %>% filter(country=="World") %>% select(sector, year, value, base, change, country)
world_oecd <- rbind(world, oecd_ch) %>% filter(sector %in% entr)

ggplot(world_oecd, aes(x=year, y=change, color=sector)) + geom_line() +
  facet_wrap(~country) + theme_stata() + theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title = "Change in GHG compared to 2000 level", x="Year", y="%") + scale_color_manual(values = c("#20639B", "#ED553B"))


```

### Korea and Germany

```{r}
entr <- c("Energy: Electricity/Heat", "Energy: Transportation")
top5_entr <- cait_ch %>% filter(country %in% c("South Korea", "Germany")) %>% filter(sector %in% entr)
ggplot(top5_entr, aes(x=year, y=change, color=sector)) + geom_line() +
  facet_wrap(~country)
```
```



# What are the main challenges in decarbonizing the energy and transport sector? 



